Bx to Rev (tentative) \thanks{Supported by organization x.}

### 2019-08-13 paper/main.tex
Update main.tex

\newcommand{\bxput}[3]{put \, [\![#1]\!] \ #2 \ #3}
\newcommand{\bxputinline}[1]{put \, [\![#1]\!]}
\newcommand{\bxget}[2]{get \, [\![#1]\!] \ #2}
\newcommand{\bxgetinline}[1]{get \, [\![#1]\!]}

\newcommand{\revput}[3]{put$_{REV}$ \ {#1} \ {#2} \ {#3}}
\newcommand{\revget}[2]{get$_{REV}$ \ {#1} \ {#2}}

@@

% Story ->
% reversible is needed not to lose any information -> pg is introduced.
% pg is not good for lassoc -> cpg (xpg2) is introduced.
% cpg can be improved -> kpg(xpg) is introduced.
% kpg is slower than pg for rassoc. combined version -> xpg.

@@

$\bxput{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$

$\bxput{Replace}{s}{v} = v$

$\bxput{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
    \tab \product{(\bxput{bx_1}{s_1}{v_1})}{(\bxput{bx_2}{s_2}{v_2})}$

$\bxput{\rearrS{f}{bx}}{s}{v} = f^{-1}(\bxput{bx}{(f \ s)}{v})$

$\bxput{\rearrV{g}{bx}}{s}{v} = \bxput{bx}{s}{(g \ v)}$

$\bxput{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s}{v} = \\
    \tab \text{if } {cond_{sv1} \ s \ v} \\
    \tab \text{then let } s' = \bxput{bx_1}{s}{v} \\
    \tab \text{else let } s' = \bxput{bx_2}{s}{v} \\
    \tab \text{fi } cond_{s1} \ s'; \text{ return } s'$

$\bxput{(bx_1 \circ bx_2)}{s}{v} = \bxput{bx_1}{s}{(\bxput{bx_2}{(\bxget{bx_1}{s})}{v})}$

@@

$\bxget{Skip \ h}{s} = h \ s$

$\bxget{Replace}{s} = s$

$\bxget{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
    \tab \product{(\bxget{bx_1}{s_1})}{(\bxget{bx_2}{s_2})}$

$\bxget{\rearrS{f}{bx}}{s} = \bxget{bx}{(f \ s)}$

$\bxget{\rearrV{g}{bx}}{s} = g^{-1} (\bxget{bx}{s})$

$\bxget{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s} =\\
    \tab \text{if } {cond_{s1} \ s}\\
    \tab \text{then let } v' = \bxget{bx_1}{s}\\
    \tab \text{else let } v' = \bxget{bx_2}{s}\\
    \tab \text{fi } {cond_{sv1} \ s \ v'}; \text{ return } v'$

$\bxget{bx_1 \circ bx_2}{s} = \bxget{bx_2}{(\bxget{bx_1}{s})}$

@@

% \noindent
% \revput{(Skip $h$)}{s}{v} = \sif{$h$ s = v}{(s, v)}{fail}

% \noindent
% \revput{Replace}{s}{v} = (v, s)

% \noindent
% \revput{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}} =\\
% \hspace{5mm}let (s1, g1) = \revput{$bx_1$}{$s_1$}{$v_1$} in\\
% \hspace{5mm}let (s2, g2) = \revput{$bx_2$}{$s_2$}{$v_2$} in\\
% \hspace{5mm}((s1, s2), (g1, g2))

% \noindent
% \revput{(\rearrS{$f$}{bx})}{s}{v} =\\
% \hspace{5mm}let (s1, g1) = \revput{bx}{($f$ s)}{v} in\\
% \hspace{5mm}($f^{-1}$ s1, g1)

% \noindent
% \revput{(\rearrV{$g$}{bx})}{s}{v} =\\
% \hspace{5mm} let (s1, g1) = \revput{bx}{s}{($g$ v)} in\\
% \hspace{5mm} (s1, g1)

% \noindent
% \revput{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s}{v} =\\
% \hspace{5mm} if {cond$_{sv1}$ s v}\\
% \hspace{5mm} then let (s', g) = \revput{bx$_1$}{s}{v} in\\
% \hspace{5mm} else let (s', g) = \revput{bx$_2$}{s}{v} in\\
% \hspace{5mm} fi cond$_{s1}$ s'; return (s', g)

% \noindent
% \revput{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
% \hspace{5mm} let (s1, g1) = \revget{bx$_1$}{s} in\\
% \hspace{5mm} let (s2, g2) = \revput{bx$_2$}{s1}{v} in\\
% \hspace{5mm} let (s3, g3) = \revput{bx$_1$}{s}{s2} in (* need to copy s *)\\
% \hspace{5mm} (s3, (g1, g2, g3))

% \vspace{5mm}
% \noindent
% \revget{(Skip $h$)}{s} = {($h$ s, s)}

% \noindent
% \revget{Replace}{s} = (s, \textcolor{red}{Nil})

% \noindent
% \revget{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}} = \\
% \hspace{5mm} let (v1, g1) = \revget{$bx_1$}{$s_1$} in\\
% \hspace{5mm} let (v2, g2) = \revget{$bx_2$}{$s_2$} in\\
% \hspace{5mm} ((v1, v2), (g1, g2))

% \noindent
% %% from here
% \revget{(\rearrS{$f$}{bx})}{s}{v} =\\
% \hspace{5mm}let (v1, g1) = \revget{bx}{($f$ s)} in\\
% \hspace{5mm}(v1, g1)

% \noindent
% \revget{(\rearrV{$g$}{bx})}{s}{v} =\\
% \hspace{5mm} let (v1, g1) = \revget{bx}{s} in\\
% \hspace{5mm} ($g^{-1}$ v1, g1)

% \noindent
% \revget{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s} =\\
% \hspace{5mm} if {cond$_{s1}$ s}\\
% \hspace{5mm} then let (v', g) = \revget{bx$_1$}{s} in\\
% \hspace{5mm} else let (v', g) = \revget{bx$_2$}{s} in\\
% \hspace{5mm} fi cond$_{sv1}$ s v'; return (v', g)

% \noindent
% \revget{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
% \hspace{5mm} let (v1, g1) = \revget{bx$_1$}{s} in\\
% \hspace{5mm} let (v2, g2) = \revget{bx$_1$}{s} in\\
% \hspace{5mm} (v2, (g1, g2))

@@

\textbf{Definition:} \\
$\pg{bx}{s}{v} = (\bxput{bx}{s}{v}, \bxget{bx}{s})$ \\

@@

Proof:\\
$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
    \tab = \pg{bx}{(\bxput{bx}{s}{v})}{(\bxget{bx}{s})} \quad [pg \text{ definition}] \\
    \tab = (put \ [\![bx]\!] \ (\bxput{bx}{s}{v}) \ (\bxget{bx}{s}), \bxget{bx}{(\bxput{bx}{s}{v})})  \quad [pg \text{ definition}] \\
    \tab = (put \ [\![bx]\!] \ (\bxput{bx}{s}{v}, \bxget{bx}{s}), v) \quad [\text{PUTGET}] \\
    \tab = (put \ [\![bx]\!] \ (s, \bxget{bx}{s}), v) \quad [\text{PUTPUT}] \\
    \tab = (s, v) \quad [\text{GETPUT}]$

@@

There is a trick in the construction of $\pginline{Skip \ h}$. The first equality is simply based on the definitions of $pg$, $\bxputinline{Skip \ h}$ and $\bxgetinline{Skip \ h}$. The second one tuples two results of $put$ and $get$ in the body of the if-expression. And the last one is quite obivious. What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\bxgetinline{Skip \ h}$.\\

@@

$\pg{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} \\
    \tab = (\product{(\bxput{bx_1}{s_1}{v_1})}{(\bxput{bx_2}{s_2}{v_2})}, \product{(\bxget{bx_1}{s_1})}{(\bxget{bx_2}{s_2})}) \\
    \tab = \text{let } (s_1, v_1) = \pg{bx_1}{s_1}{v_1} \text{ in} \\
        \qtab \text{let } (s_2, v_2) = \pg{bx_2}{s_2}{v_2} \text{ in} \\
        \qtab (\product{s_1}{s_2}, \product{v_1}{v_2})$

$\pg{\rearrS{f}{bx}}{s}{v} \\
    \tab = (f^{-1}(\bxput{bx}{(f \ s)}{v}), \bxget{bx}{(f \ s)}) \\
    \tab = \text{let } (s, v) = \pg{bx}{(f \ s)}{v} \text{ in} \\
        \qtab (f^{-1} s, v)$

$\pg{\rearrV{g}{bx}}{s}{v} \\
    \tab = (\bxput{bx}{s}{(g v)}, g^{-1} (\bxget{bx}{s})) \\
    \tab = \text{let } (s, v) = \pg{bx}{s}{(g v)} \text{ in} \\
        \qtab (s, g^{-1} v)$\\

@@

$\pg{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s}{v} \\
    \tab = (\text{if } {cond_{sv1} \ s \ v} \\
    \qtab \text{then let } s' = \bxput{bx_1}{s}{v} \\
    \qtab \text{else let } s' = \bxput{bx_2}{s}{v} \\
    \qtab \text{fi } cond_{s1} \ s'; \text{ return } s', \\
    \qtab \quad \text{if } {cond_{s1} \ s}\\
    \qtab \quad \text{then let } v' = \bxget{bx_1}{s}\\
    \qtab \quad \text{else let } v' = \bxget{bx_2}{s}\\
    \qtab \quad \text{fi } {cond_{sv1} \ s \ v'}; \text{ return } v') \\
    \tab = \text{if } {cond_{sv1} \ s \ v} \ and \ {cond_{s1} \ s} \\
        \qtab \text{then let } (s', v') = \pg{bx_1}{s}{v}\\
        \qtab \text{else let } (s', v') = \pg{bx_2}{s}{v}\\
        \qtab \text{fi } cond_{v1} \ s' \ and \ cond_{sv1} \ s \ v'; \text{ return } (s',v')$\\

A restriction for $\pginline{case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\bxputinline{case}$ as well as $\bxgetinline{case}$. If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions. This means that two entering conditions of $\bxputinline{case}$ and $\bxgetinline{case}$ are not always simultaneously satified. The evaluated BXs are different in the $put$ and $get$ directions for combinations $((cond_{sv1} \ s \ v) \ and \ (not (cond_{s1} \ s)))$ and $((not (cond_{sv1} \ s \ v) \ and \ (cond_{s1} \ s))$, which are restricted in this paper. This does not happen for the others, which the construction of $\pginline{case}$ uses above. \\

$\pg{bx_1 \circ bx_2}{s}{v} \\
    \hspace*{5mm} = (\bxput{bx_1}{s}{(\bxput{bx_2}{(\bxget{bx_1}{s})}{v})}, \bxget{bx_2}{(\bxget{bx_1}{s})}) \\
    \hspace*{5mm} = \text{let } a = \bxget{bx_1}{s} \text{ in} \\
        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in} \\
        \hspace*{5mm} \ \quad (d, c) \\
    \hspace*{5mm} = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in} \\
        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in} \\
        \hspace*{5mm} \ \quad (d, c) \\
    \hspace*{5mm} = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in} \\
        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{\textcolor{red}{f}}{b} \text{ in} \\
        \hspace*{5mm} \ \quad (d, c)$\\

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $(\text{let } a = \bxget{bx_1}{s} \text{ in})$ into $(\text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in})$. It is possible if we consider $(\bxget{bx_1}{s})$ as the second element of $(\pg{bx_1}{s}{dummy})$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ direction is valid. The last equality substitutes $s$ in $(\text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in})$ by $f$ which is the evaluated result of $\bxput{bx_1}{s}{dummy}$. This transformation is feasible due to the PUTPUT law. Then, $e$ in the result pair $(d,e)$ of the new let-expression equals $dummy$. So we can realize that there is no loss information when computing a composition.

@@

$fst = \lambda (Con (x1, x2)). x1$\\
$snd = \lambda (Con (x1, x2)). x2$\\
$con = \lambda ks1. \lambda ks2. \lambda x. Con(ks1 \ x, ks2 \ x)$\\

$\cpg{bx}{ks}{kv}{s}{v}$ \\
    \tab $ks, kv$ evaluate actual values\\
    \tab $s, v$ keep evaluated values\\
\vspace{3mm}

@@

$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
    \tab (ks1, kv1, s1, v1) \Leftarrow \cpg{bx1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
    \tab (ks2, kv2, s2, v2) \Leftarrow \cpg{bx2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
    \qtab ((con \ ks1 \ ks2, con \ kv1 \ kv2, Con \ s1 \ s2, Con \ v1 \ v2)$\\

$\cpg{RearrS \ f1 \ f2 \ bx}{ks}{kv}{s}{v} =\\
    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f1 \circ ks}{kv}{f1 \ s}{v};\\
    \qtab (f2 \circ ks, kv, s, v)$\\

$\cpg{RearrV \ g1 \ g2 \ bx}{ks}{kv}{s}{v} =\\
    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g1 \circ kv}{s}{g1 \ v};\\
    \qtab (ks, g2 \circ kv, ks, g2 \ v)$\\

$\cpg{Case \ c \ bx1 \ bx2}{ks}{kv}{s}{v} =\\
    \tab \text{if} \ c \ s \ v \ \text{then}\\
        \qtab \cpg{bx1}{ks}{kv}{s}{v}\\
    \tab \text{else} \\ 
        \qtab \cpg{bx2}{ks}{kv}{s}{v}$\\

$\cpg{bx1 \circ bx2}{ks}{kv}{s}{v} =\\
    \tab (ks1, kv1, \underline{s1}, v1) \Leftarrow \cpg{bx1}{ks}{id}{s}{const$\_$dummy \ v};\\
    \tab (ks2, kv2, s2, v2) \Leftarrow \cpg{bx2}{kv1}{kv}{v1}{v};\\
        \qtab (ks1 \circ ks2, kv2,  ks1 \ s2, v2)$\\

\begin{itemize}
\item This is faster than pg for left assoc BX programs.
\item (Slower for right assoc programs)
\item $s1'$ in composition is redundant.  
\item Next, we will extend this to avoid this redundant computation.
\end{itemize}

@@

$\kpg{bx}{ks'}{kv'}{ks}{kv}{s}{v}$ \\
    \tab $ks', kv'$ keep modification information\\
    \tab $ks, kv$ evaluate actual values\\
    \tab $s, v$ keep evaluated values\\

$\kpg{Skip \ h}{ks'}{kv'}{ks}{kv}{s}{v} =\\
    \tab s \Leftarrow ks \ s;\\
    \tab v \Leftarrow kv \ v;\\
    \tab \text{if} \ h \ s = v \ \text{then} \ (ks', kv', id, id, s, v) \ \text{else fail}$\\

$\kpg{Replace}{ks'}{kv'}{ks}{kv}{s}{v} = (kv', ks', kv, ks, v, s)$\\

>>>>

\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}

\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}

@@

% Story ->
% reversible is needed not to lose any information -> pg is introduced.
% pg is not good for lassoc -> cpg (xpg_2) is introduced.
% cpg can be improved -> kpg(xpg) is introduced.
% kpg is slower than pg for rassoc. combined version -> xpg.

@@

$\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$

$\putbx{Replace}{s}{v} = v$

$\putbx{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
    \tab \product{(\putbx{bx_1}{s_1}{v_1})}{(\putbx{bx_2}{s_2}{v_2})}$

$\putbx{\rearrS{f}{bx}}{s}{v} = f^{-1}(\putbx{bx}{(f \ s)}{v})$

$\putbx{\rearrV{g}{bx}}{s}{v} = \putbx{bx}{s}{(g \ v)}$

$\putbx{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} = \\
    \tab \text{if } {cond_{sv_1} \ s \ v} \\
    \tab \text{then let } s' = \putbx{bx_1}{s}{v} \\
    \tab \text{else let } s' = \putbx{bx_2}{s}{v} \\
    \tab \text{fi } cond_{s_1} \ s'; \ \text{return} \ s'$

$\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$

@@

$\getbx{Skip \ h}{s} = h \ s$

$\getbx{Replace}{s} = s$

$\getbx{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
    \tab \product{(\getbx{bx_1}{s_1})}{(\getbx{bx_2}{s_2})}$

$\getbx{\rearrS{f}{bx}}{s} = \getbx{bx}{(f \ s)}$

$\getbx{\rearrV{g}{bx}}{s} = g^{-1} (\getbx{bx}{s})$

$\getbx{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s} =\\
    \tab \text{if} \ {cond_{s_1} \ s}\\
    \tab \text{then let} \ v' = \getbx{bx_1}{s}\\
    \tab \text{else let} \ v' = \getbx{bx_2}{s}\\
    \tab \text{fi} \ {cond_{sv_1} \ s \ v'}; \ \text{return} \ v'$

$\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$

@@

$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$

$\putrev{Replace}{s}{v} = (v, s)$

$\putrev{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
    \tab \text{let} \ (s_1, g_1) = \putrev{bx_1}{s_1}{v_1} \ \text{in}\\
    \tab \text{let} \ (s_2, g_2) = \putrev{bx_2}{s_2}{v_2} \ \text{in}\\
        \qtab ((s_1, s_2), (g_1, g_2))$

$\putrev{\rearrS{f}{bx}}{s}{v} =\\
    \tab \text{let} \ (s_1, g_1) = \putrev{bx}{(f \ s)}{v} \ \text{in}\\
        \qtab (f^{-1} s_1, g_1)$

$\putrev{\rearrV{g}{bx}}{s}{v} =\\
    \tab \text{let} \ (s_1, g_1) = \putrev{bx}{s}{(g \ v)} \ \text{in}\\
        \qtab (s_1, g_1)$

$\putrev{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} =\\
    \tab \text{if} \ cond_{sv_1} \ s \ v\\
    \tab \text{then let} \ (s', g) = \putrev{bx_1}{s}{v} \ \text{in}\\
    \tab \text{else let} \ (s', g) = \putrev{bx_2}{s}{v} \ \text{in}\\
    \tab \text{fi} \ cond_{s_1} \ s'; \ \text{return} \ (s', g)$

$\putrev{bx_1 \circ bx_2}{s}{v} =\\
    \tab \text{let} \ (s_1, g_1) = \getrev{bx_1}{s} \ \text{in}\\
    \tab \text{let} \ (s_2, g_2) = \putrev{bx_2}{s_1}{v} \ \text{in}\\
    \tab \text{let} \ (s_3, g_3) = \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
        \qtab (s_3, (g_1, g_2, g_3))$

$\getrev{Skip \ h}{s} = {(h \ s, s)}$

$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$

$\getrev{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
    \tab \text{let} \ (v_1, g_1) = \getrev{bx_1}{s_1} \ \text{in}\\
    \tab \text{let} \ (v_2, g_2) = \getrev{bx_2}{s_2} \ \text{in}\\
        \qtab ((v_1, v_2), (g_1, g_2))$

$\getrev{\rearrS{f}{bx}}{s}{v} =\\
    \tab \text{let} \ (v_1, g_1) = \getrev{bx}{(f \ s)} \ \text{in}\\
        \qtab (v_1, g_1)$

$\getrev{\rearrV{g}{bx}}{s}{v} =\\
    \tab \text{let} \ (v_1, g_1) = \getrev{bx}{s} \ \text{in}\\
        \qtab (g^{-1} v_1, g_1)$

$\getrev{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s} =\\
    \tab \text{if} \ cond_{s_1} \ s\\
    \tab \text{then let} \ (v', g) = \getrev{bx_1}{s} \ \text{in}\\
    \tab \text{else let} \ (v', g) = \getrev{bx_2}{s} \ \text{in}\\
    \tab \text{fi} \ cond_{sv_1} \ s \ v'; \ \text{return} \ (v', g)$

$\getrev{(bx_1 \circ bx_2)}{s}{v} =\\
    \tab let (v_1, g_1) = \getrev{bx_1}{s} in\\
    \tab let (v_2, g_2) = \getrev{bx_1}{s} in\\
        \qtab (v_2, (g_1, g_2))$

@@

\textbf{Definition:} \\
$\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$ \\

@@

Proof:\\
$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
    \tab = (s, v) \quad [\text{GETPUT}]$

@@

There is a trick in the construction of $\pginline{Skip \ h}$. The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$. The second one tuples two results of $put$ and $get$ in the body of the if-expression. And the last one is quite obivious. What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\

@@

$\pg{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} \\
    \tab = (\product{(\putbx{bx_1}{s_1}{v_1})}{(\putbx{bx_2}{s_2}{v_2})}, \product{(\getbx{bx_1}{s_1})}{(\getbx{bx_2}{s_2})}) \\
    \tab = \text{let} \ (s_1, v_1) = \pg{bx_1}{s_1}{v_1} \ \text{in} \\
        \qtab \text{let} \ (s_2, v_2) = \pg{bx_2}{s_2}{v_2} \ \text{in} \\
        \qtab (\product{s_1}{s_2}, \product{v_1}{v_2})$

$\pg{\rearrS{f}{bx}}{s}{v} \\
    \tab = (f^{-1}(\putbx{bx}{(f \ s)}{v}), \getbx{bx}{(f \ s)}) \\
    \tab = \text{let} \ (s, v) = \pg{bx}{f \ s}{v} \ \text{in} \\
        \qtab (f^{-1} s, v)$

$\pg{\rearrV{g}{bx}}{s}{v} \\
    \tab = (\putbx{bx}{s}{(g v)}, g^{-1} (\getbx{bx}{s})) \\
    \tab = \text{let} \ (s, v) = \pg{bx}{s}{g \ v} \ \text{in} \\
        \qtab (s, g^{-1} v)$\\

@@

$\pg{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} \\
    \tab = (\text{if} \ {cond_{sv_1} \ s \ v} \\
    \qtab \text{then let} \ s' = \putbx{bx_1}{s}{v} \\
    \qtab \text{else let} \ s' = \putbx{bx_2}{s}{v} \\
    \qtab \text{fi} \ cond_{s_1} \ s'; \text{ return } s', \\
    \qtab \quad \text{if} \ {cond_{s_1} \ s}\\
    \qtab \quad \text{then let} \ v' = \getbx{bx_1}{s}\\
    \qtab \quad \text{else let} \ v' = \getbx{bx_2}{s}\\
    \qtab \quad \text{fi} \ {cond_{sv_1} \ s \ v'}; \text{ return } v') \\
    \tab = \text{if} \ {cond_{sv_1} \ s \ v} \ and \ {cond_{s_1} \ s} \\
        \qtab \text{then let} \ (s', v') = \pg{bx_1}{s}{v}\\
        \qtab \text{else let} \ (s', v') = \pg{bx_2}{s}{v}\\
        \qtab \text{fi} \ cond_{v_1} \ s' \ and \ cond_{sv_1} \ s \ v'; \text{ return } (s',v')$\\

A restriction for $\pginline{case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{case}$ as well as $\getbxinline{case}$. If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions. This means that two entering conditions of $\putbxinline{case}$ and $\getbxinline{case}$ are not always simultaneously satified. The evaluated BXs are different in the $put$ and $get$ directions for combinations $((cond_{sv_1} \ s \ v) \ and \ (not (cond_{s_1} \ s)))$ and $((not (cond_{sv_1} \ s \ v) \ and \ (cond_{s_1} \ s))$, which are restricted in this paper. This does not happen for the others, which the construction of $\pginline{case}$ uses above. \\

$\pg{bx_1 \circ bx_2}{s}{v} \\
    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tab = \text{let} \ a = \getbx{bx_1}{s} \ \text{in} \\
        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
        \qtab \text{let} \ (d, e) = \pg{bx_1}{s}{b} \ \text{in} \\
        \qtab (d, c) \\
    \tab = \text{let} \ (f, a) = \pg{bx_1}{s}{dummy} \ \text{in} \\
        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
        \qtab \text{let} \ (d, e) = \pg{bx_1}{s}{b} \ \text{in} \\
        \qtab (d, c) \\
    \tab = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \ \text{in} \\
        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
        \qtab \text{let} \ (d, e) = \pg{bx_1}{\textcolor{red}{f}}{b} \ \text{in} \\
        \qtab (d, c)$\\

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $(\text{let } a = \getbx{bx_1}{s} \text{ in})$ into $(\text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in})$. It is possible if we consider $(\getbx{bx_1}{s})$ as the second element of $(\pg{bx_1}{s}{dummy})$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ direction is valid. The last equality substitutes $s$ in $(\text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in})$ by $f$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$. This transformation is feasible due to the PUTPUT law. Then, $e$ in the result pair $(d,e)$ of the new let-expression equals $dummy$. So we can realize that there is no loss information when computing a composition.

@@

$fst = \lambda (Con (x_1, x_2)). x_1$\\
$snd = \lambda (Con (x_1, x_2)). x_2$\\
$con = \lambda ks_1. \lambda ks_2. \lambda x. Con(ks_1 \ x, ks_2 \ x)$\\

$\cpg{bx}{ks}{kv}{s}{v}$ \\
    \tab $ks, kv$ keep modification information\\
    \tab $s, v$ keep evaluated values\\
\vspace{3mm}

@@

$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, Con \ s_1 \ s_2, Con \ v_1 \ v_2)$\\

$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
    \qtab (f_2 \circ ks, kv, s, v)$\\

$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\

$\cpg{Case \ c \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
    \tab \text{if} \ c \ s \ v \ \text{then}\\
        \qtab \cpg{bx_1}{ks}{kv}{s}{v}\\
    \tab \text{else} \\ 
        \qtab \cpg{bx_2}{ks}{kv}{s}{v}$\\

$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
    \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{const$\_$dummy \ v};\\
    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$\\

\begin{itemize}
\item This is faster than pg for left assoc BX programs.
\item (Slower for right assoc programs)
\item $s_1'$ in composition is redundant.  
\item Next, we will extend this to avoid this redundant computation.
\end{itemize}

@@

$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$ \\
    \tab $ks, kv$ keep modification information\\
    \tab $ks', kv'$ evaluate actual values\\
    \tab $s, v$ keep evaluated values\\

$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\

$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
